---
title: "Notebook 2: Part 2"
output: html_notebook
---

# Health Burden

In this first code section, load your data and any necessary libraries (`dplyr`, `tidyr`, `ggplot`, any others you like).
```{r, echo=F}
# Set up (install packages if you don't have them)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
risk.data <- read.csv('./data/prepped/risk-data.csv', stringsAsFactors = FALSE) 
```

## Data Structure
To get a basic sense of your dataset, check the following:

- What is the size of the dataset (rows, columns)?
- How many countries are present in the dataset?
- What are the age-groups in the dataset?
- What are the variables (columns) present in the dataset?
- What is the data type of each variable?
- What are the summary statistics of each quantitative variable (_hint: `summary()`_)

```{r}
# Code goes here
```

## Univariate Analysis
To explore the distribution of each risk-column in a visual way, create a histogram of each risk-column of interest (e.g., `alcohol_use`, `drug_use`, etc.). For easier comparisons, you may want to store each histogram in a variable, then render all of the charts using the `grid.arrange` function from the `gridExtra` package.

If you want to do this in an more advanced (and concise!) `R` way, you should **reshape your data** using the `gather` function to combine the risks (which are currently each their own column) into two columns (one for the _name_ of the risk, and the other for the _value_ of the death rate of that risk). Then, you can use the `facet_wrap` method of the ggplot library.

- Note what you find surprising!


```{r}
# Make histograms for each risk column

```

_list some observations from your histograms here using Markdown code_:

- Observation one
- Observation two
- etc...

## Univariate analysis (by age)
In this section, you should **write a function** to investigate how a given risk (column) varies by **age group** (the risk -- column name -- will be the argument to your function). More specifically, the function should create a set of histograms (one for each age-group) to show the distribution of death rates for a given risk factor. You should use the `facet_wrap` function (from `ggplot2`) to create the multi-plot layout.

```{r}
# Write your function here


# Create a plot to demonstrate that it works

```

_list some observations from your histograms here using Markdown code_:

- Observation one
- Observation two
- etc...

## Univariate analysis (by sex)
In this section, you should investigate how a given risk (column) varies by **sex**. To do this, **write a function** to create a single histogram with _overlapping distributions_ for males and females (for a given risk factor). 

_Hint_: set the `fill` of your `ggplot()` aesthetic, and turn down the _alpha_ (opacity) to help reduce occlusion in the chart. 

```{r}
# Write your function here


# Create plot to demonstrate that it works

```

_list some observations from your histograms here using Markdown code_:

- Observation one
- Observation two
- etc...


## Bivariate analysis
As a final step, create a **scatterplot matrix** of all of your risks to visually assess the correlation across death rates.

```{r}
# Create your scatterplot matrix here
```

_list some observations from your histograms here using Markdown code_:

- Observation one
- Observation two
- etc...

When you've finished this step, compile your code to an `html` document. 